---
title: "03_analysis"
author: "Joey O'Brien"
date: "11/5/2021"
output: html_document
---

When using Spark with R the data is stored in the former. We send commands from `R` to Spark which performs analysis within its own system before sending the results only back to `R`.

Let's start by loading a dataset into Spark as before

```{r}
cars <- copy_to(sc, mtcars)
```

`R` sends syntax to interact with the SQL side of Spark so for example running

```{r}
summarize_all(cars, mean)
```
makes use of the following SQL query

```{r}
summarize_all(cars, mean) %>%
  show_query()
```

and can do more advanced queries using the `tidyverse` language for example

```{r}
cars %>%
  mutate(transmission = ifelse(am == 0, "automatic", "manual")) %>%
  group_by(transmission) %>%
  summarise_all(mean)
```

Of course there are some functions built directly into Spark e.g., the _ml_corr_ function used to calculate correlations
```{r}
ml_corr(cars)
```
but even then we can just use `R` functionality which which then runs _ml_corr_ in the background after the translation in `sparklyr`

```{r}
library(corrr)
correlate(cars, use = "pairwise.complete.obs", method = "pearson") 
```
### Visualizations

In `sparklyr` the plotting is performed by all the analysis taking place in Spark and the final plotting part taking place in `R`

```{r}
library(ggplot2)
ggplot(aes(as.factor(cyl), mpg), data = mtcars) + geom_col()
```
we can also do some data wrangling prior to plotting but we have to collect the wrangled data in `R` prior to plotting

```{r}
car_group <- cars %>%
  group_by(cyl) %>%
  summarise(mpg = sum(mpg, na.rm = TRUE)) %>%
  collect() %>%
  print()
```

```{r}
ggplot(aes(as.factor(cyl), mpg), data = car_group) + 
  geom_col(fill = "#999999") + coord_flip()
```
Alternatively you can use the `dbplot` package which is built for Spark and can perform both the wrangling and plotting at once

```{r}
library(dbplot)

cars %>%
dbplot_histogram(mpg, binwidth = 3) +
labs(title = "MPG Distribution",
     subtitle = "Histogram over miles per gallon")
```
### Modelling

Just like plotting most modelling in `sparklyr` is performed just like it would be in normal `R` and we'll see more in the next set of notes. Here's just a flavour.

```{r}
cars %>% 
  ml_linear_regression(mpg ~ .) %>%
  summary()
```

```{r}
cars %>% 
  ml_linear_regression(mpg ~ hp + cyl) %>%
  summary()
```

