---
title: "04_modelling"
author: "Joey O'Brien"
date: "11/5/2021"
output: html_document
---

To highlight the machine learning functionality in `spraklyr` he first load in the `okCupid` dataset which consists of 60,000 user profiles from an online dating site.

```{r}
download.file(
  "https://github.com/r-spark/okcupid/raw/master/profiles.csv.zip",
  "okcupid.zip")

unzip("okcupid.zip", exdir = "data")
unlink("okcupid.zip")
```

We will try to answer the following question

_Predict whether someone is actively working—that is, not retired, a student, or unemployed._

### exploratory data analysis

Let's load the data into Spark straight away using the `spark_read_csv` function

```{r}
library(sparklyr)
library(ggplot2)
library(dbplot)
library(dplyr)

sc <- spark_connect(master = "local", version = "2.3")

okc <- spark_read_csv(
  sc, 
  "data/profiles.csv", 
  escape = "\"", 
  memory = FALSE,
  options = list(multiline = TRUE)
) %>%
  mutate(
    height = as.numeric(height),
    income = ifelse(income == "-1", NA, as.numeric(income))
  ) %>%
  mutate(sex = ifelse(is.na(sex), "missing", sex)) %>%
  mutate(drinks = ifelse(is.na(drinks), "missing", drinks)) %>%
  mutate(drugs = ifelse(is.na(drugs), "missing", drugs)) %>%
  mutate(job = ifelse(is.na(job), "missing", job))
```
```{r}
glimpse(okc)
```
and let's add the response variable we want to use which is about working or not

```{r}
okc <- okc %>%
  mutate(
    not_working = ifelse(job %in% c("student", "unemployed", "retired"), 1 , 0)
  )

okc %>% 
  group_by(not_working) %>% 
  tally()
```

Let's begin by splitting the dataset into a test and training subset, we could do this a nubmer of ways but let's do the most simple which is just a random split

```{r}
data_splits <- sdf_random_split(okc, training = 0.8, testing = 0.2, seed = 42)
okc_train <- data_splits$training
okc_test <- data_splits$testing
```

and we can check if there is a class imbalance in out training dataset
```{r}
okc_train %>%
  group_by(not_working) %>%
  tally() %>%
  mutate(frac = n / sum(n))
```

and we can do some initial summary statistics for certain columns using the `sdf_describe` function
```{r}
sdf_describe(okc_train, cols = c("age", "income"))
```

and plot some variables to get a feel for the data

```{r}
dbplot_histogram(okc_train, age)
```

_(hmmm kind of heavy tailed...)_

given we are interested in whether the individual is working or not we can look at how this quantity relates to other variables e.g., their religion

```{r}
prop_data <- okc_train %>%
  mutate(religion = regexp_extract(religion, "^\\\\w+", 0)) %>% 
  group_by(religion, not_working) %>%
  tally() %>%
  group_by(religion) %>%
  summarize(
    count = sum(n),
    prop = sum(not_working * n) / sum(n)
  ) %>%
  mutate(se = sqrt(prop * (1 - prop) / count)) %>%
  collect()

prop_data
```

and we can construct a 95% confidence interval about this proportion estimate to get an idea of religions which are more common to be employed/unemployed 

```{r}
prop_data %>%
  ggplot(aes(x = religion, y = prop)) + geom_point(size = 2) +
  geom_errorbar(aes(ymin = prop - 1.96 * se, ymax = prop + 1.96 * se),
                width = .1) +
  geom_hline(yintercept = sum(prop_data$prop * prop_data$count) /
                              sum(prop_data$count)) +
  labs(y = 'Proportion Unemployed', x = 'Religion')
```

